Future challenges for Research Software Engineering

### 2025-09-23 future_challenges_for_RSEng.tex
Mli 1 (#65)
* sort out box 1 of issue 52

* resolve box 2

* sort out Comment-3

* sort out Comment-5

* sort out Box-6

* sort out Comment-8

* sort out Comment-9

* sort out Comment-10

* sort out Comment-11


Co-authored-by: Jeremy Cohen <jcohen02@users.noreply.github.com>

Co-authored-by: Magnus Hagdorn <Magnus.Hagdorn@marsupium.org>

* take care of jeremys comment

* work out today's idea

---------

Co-authored-by: Florian Goth <florian.goth@uni-wuerzburg.de>
Co-authored-by: Jeremy Cohen <jcohen02@users.noreply.github.com>
Co-authored-by: Magnus Hagdorn <Magnus.Hagdorn@marsupium.org>

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\authorOrcid{0000-0003-2707-4790}\authorRef{1},
Leyla Jael Castro\authorOrcid{0000-0003-3986-0510}\authorRef{2},
Gerasimos Chourdakis\authorOrcid{0000-0002-3977-1385}\authorRef{3},
\texorpdfstring{\\}{} % line break
Simon Christ\authorOrcid{0000-0002-5866-1472}\authorRef{4},
Jeremy Cohen\authorOrcid{0000-0003-4312-2537}\authorRef{5},
Nicholas A. Del Grosso\thinspace\authorRef{6},
\texorpdfstring{\\}{} % line break
Jean-Noël Grad\authorOrcid{0000-0002-5821-4912}\authorRef{7},
Magnus Hagdorn\authorOrcid{0000-0002-5076-4864}\authorRef{8},
Toby Hodges\authorOrcid{0000-0003-1766-456X}\authorRef{9},
Jan Linxweiler\authorOrcid{0000-0002-2755-5087}\authorRef{10},
\texorpdfstring{\\}{} % line break
Frank Löffler\authorOrcid{0000-0001-6643-6323}\authorRef{11},
Michele Martone\authorOrcid{0000-0003-3239-8554}\authorRef{12},
Jan Philipp Thiele\authorOrcid{0000-0002-8901-6660}\authorRef{10}\textsuperscript{,}\authorRef{13}
} % Authors and references to addresses
\institute{% Institutes with labels
\autlabel{1} Institut für theoretische Physik 1, University of Würzburg, 97074, Würzburg, Germany\\
\autlabel{2} ZB MED Information Centre for Life Sciences, Cologne, Germany\\
\autlabel{3} Institute for Parallel and Distributed Systems, University of Stuttgart, Stuttgart, Germany\\
\autlabel{4} Leibniz University Hannover, Department of Cell Biology and Biophysics, Computational Biology, Germany\\
\autlabel{5} Imperial College London, London, UK\\
\autlabel{6} Institute for Experimental Epileptology and Cognition Research, Uniklinikum Bonn, Germany\\
\autlabel{7} Institute for Computational Physics, University of Stuttgart, Germany\\
\autlabel{8} Geschäftsbereich IT, Charité Universitätsmedizin Berlin, Germany\\
\autlabel{9} The Carpentries, USA\\
\autlabel{10} Technische Universität Braunschweig, Germany\\
\autlabel{11} Michael Stifel Center Jena {\&} Friedrich Schiller University Jena, Germany\\
\autlabel{12} Independent researcher, Germany\\
\autlabel{13} Weierstrass Institute, Berlin, Germany;
              Leibniz University Hannover, Institute of Applied Mathematics, Scientific Computing, Hannover, Germany\\
}
\abstract{
It is 2025 and the research software community is, admittedly, still working on establishing and growing Research Software Engineering (RSEng)
as a domain in its own right, with the hope of ensuring that it gets the recognition it deserves.
As the field develops, the role and value of Research Software Engineers (RSEs) becomes clearer.
With the growing importance and scope of software engineering in research,
researchers who code now begin to fill dedicated, specialized RSE roles crucial for the sustainable development
of long-term research software projects that can rapidly innovate.
Looking into the future, participants of the deRSE25 conference discussed how RSEng could evolve in response to the
rapid changes in the broader social, technological, and organisational landscape.
This paper presents the main discussion points of the conference participants,
imagining a future in which RSEs have a well-defined and well-integrated role in research institutions,
working as engineers of ever more complex research software projects,
taming a growing range of methods and tools,
and guiding researchers through technical and ethical considerations
related to research software.
} % Abstract of the article
\keywords{Future RSE, Evolving field, Vision} % Keywords for the article

>>>>

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\authorOrcid{0000-0003-2707-4790}\authorRef{1},
Leyla Jael Castro\authorOrcid{0000-0003-3986-0510}\authorRef{2},
Gerasimos Chourdakis\authorOrcid{0000-0002-3977-1385}\authorRef{3},
Simon Christ\authorOrcid{0000-0002-5866-1472}\authorRef{4},
\texorpdfstring{\\}{} % line break
Jeremy Cohen\authorOrcid{0000-0003-4312-2537}\authorRef{5},
Nicholas A. Del Grosso\thinspace\authorRef{6},
Jean-Noël Grad\authorOrcid{0000-0002-5821-4912}\authorRef{7},
\texorpdfstring{\\}{} % line break
Magnus Hagdorn\authorOrcid{0000-0002-5076-4864}\authorRef{8},
Toby Hodges\authorOrcid{0000-0003-1766-456X}\authorRef{9},
Jan Linxweiler\authorOrcid{0000-0002-2755-5087}\authorRef{10}\textsuperscript{,}\authorRef{11},
\texorpdfstring{\\}{} % line break
Frank Löffler\authorOrcid{0000-0001-6643-6323}\authorRef{12}\textsuperscript{,}\authorRef{11},
Michele Martone\authorOrcid{0000-0003-3239-8554}\authorRef{13},
Jan Philipp Thiele\authorOrcid{0000-0002-8901-6660}\authorRef{10}\textsuperscript{,}\authorRef{14}
} % Authors and references to addresses
\institute{% Institutes with labels
\autlabel{1} Institut für theoretische Physik 1, University of Würzburg, 97074, Würzburg, Germany\\
\autlabel{2} ZB MED Information Centre for Life Sciences, Cologne, Germany\\
\autlabel{3} Institute for Parallel and Distributed Systems, University of Stuttgart, Stuttgart, Germany\\
\autlabel{4} Leibniz University Hannover, Department of Cell Biology and Biophysics, Computational Biology, Germany\\
\autlabel{5} Imperial College London, London, UK\\
\autlabel{6} Institute for Experimental Epileptology and Cognition Research, Uniklinikum Bonn, Germany\\
\autlabel{7} Institute for Computational Physics, University of Stuttgart, Germany\\
\autlabel{8} Geschäftsbereich IT, Charité Universitätsmedizin Berlin, Germany\\
\autlabel{9} The Carpentries, USA\\
\autlabel{10} Technische Universität Braunschweig, Germany\\
\autlabel{11} de-RSE e.V.---Society for Research Software in Germany\\
\autlabel{12} Competence Center Digital Research, Friedrich Schiller University Jena, Germany\\
\autlabel{13} Independent researcher, Germany\\
\autlabel{14} Weierstrass Institute, Berlin, Germany;
              Leibniz University Hannover, Institute of Applied Mathematics, Scientific Computing, Hannover, Germany\\
}
\abstract{
It is 2025 and the research software community is, admittedly, still working on establishing and growing Research Software Engineering (RSEng)
as a domain in its own right, with the hope of ensuring that it gets the recognition it deserves.
As the field develops, the role and value of Research Software Engineers (RSEs) becomes clearer.
With the growing importance and scope of software engineering in research,
researchers who code now begin to fill dedicated, specialized RSE roles crucial for the sustainable development
of long-term research software projects that can rapidly innovate.
Looking into the future, participants of the deRSE25 conference discussed how RSEng could evolve in response to the
rapid changes in the broader social, technological, and organisational landscape.
This paper presents the main discussion points of the conference participants,
imagining a future in which RSEs have a well-defined and well-integrated role in research institutions,
working as engineers of ever more complex research software projects,
taming a growing range of methods and tools,
and guiding researchers through technical and ethical considerations
related to research software.
} % Abstract of the article
\keywords{Future RSE, Evolving field, Vision} % Keywords for the article
